Modular Model-Checking of a Byzantine Fault-Tolerant Protocol

### 2017-03-01 paper/paper.tex
Remove redundant institute footnote marker
final fixes from Lee.

\usepackage{alltt}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{ifthen}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{tikz-cd}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{xspace}

@@

\author{Benjamin F Jones\inst{1} \and Lee Pike\inst{1}}

@@

\begin{abstract}
With proof techniques like IC3 and $k$-induction, model-checking scales further
than ever before. Still, fault-tolerant distributed systems are particularly
challenging to model-check given their large state spaces and
non-determinism. The typical approach to controlling complexity is to construct
ad-hoc abstractions of faults, message-passing, and behaviors. However, these
abstractions come at the price of divorcing the model from its implementation
and making refactoring difficult. In this work, we present a model for
fault-tolerant distributed system verification that combines ideas from the
literature including calendar automata, symbolic fault injection, and abstract
transition systems, and then use it to model-checking various implementations of
the Hybrid Oral Messages algorithm that differ in the fault model, timing model,
and local node behavior. We show that despite being implementation-level models,
the verifications are scalable and modular, insofar as isolated changes to an
implementation require isolated changes to the model and proofs. This work is
carried out in SAL model-checker.
\end{abstract}

@@

In this paper, we present a fault-tolerant distributed systems model, and use that model to verify several variant implementations of the Byzantine fault-tolerant Hybrid Oral Messages algorithm (\OMH)~\cite{Lincoln-Rushby}. The model combines various ideas from the literature and individual verification efforts to build scalable and modular formal models suitable for infinite-state model-checking, and it reduces the need for ad-hoc abstractions and optimizations. In Section~\ref{sec:model}, we present the important aspects of the model, including calendar automata, originally developed by Dutertre and Sorea~\cite{cal}, symbolic fault-injection, and abstract transition systems for verification.

@@

Let a transition system $\mathcal{M} = (S, I, \rightarrow)$, be a set of states $S$, a set of initial states $I \subseteq S$, and a transition relation $\rightarrow \subseteq S \times S$. We implicitly assume a set of state variables such that each state $\sigma \in S$ is a total function that maps state variables to values. We sometimes prime a state to denote that satisfies the transition relation: $\sigma \rightarrow \sigma'$. We also sometimes use a variable assignment notation to describe what state variables are specifically updated: e.g., $\sigma' = \sigma[v := v+1]$.

@@

\item \label{cal:d} In all states, if the current time equals a timeout, then the only transitions enabled are calendar event updates associated with the timeout: $\forall \sigma \in S$, $\exists (e_i, t_i) \in \sigma(cal)$ such that $\sigma(now) = t_i$ implies $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma'(now) = \sigma(now)$, $\sigma'(c_j) = \sigma(c_j)$ for all $c_j \in \sigma(cal)$ such that $c_j \neq c_i$, and $c_i \notin \sigma'(cal)$.

@@

The Oral Messages algorithm and its variants and its variants have a long history of formal verification. \OM{1} was verified in both the PVS and ACL2 interactive theorem-provers~\cite{Young97:IC}. Also in ACL2, an implementation of a circuit design to implement \OM{1} is given~\cite{om-acl2-impl}; the low-level model most closely relates to our level of detail. A refinement-based verification approach is used, and \OM{1} is specialized to fixed number of nodes. Bokor~\emph{et~al.} describe a message-passing model for synchronous distributed algorithms that is particularly amenable to partial-order reduction for explicit-state model-checking~\cite{Bokor2010}. The model is efficient for up to five nodes, but results are not presented beyond that. Very recently, Jovanovi{\'{c}} and Dutertre use a ``flattened'' high-level model of \OM{1} as a benchmark for IC3 augmented with $k$-induction~\cite{pdr-kind}.

@@

% ------------------------------------------------------------
\section{Conclusions}\label{sec:conclusions}
This work fits within a larger project, in collaboration with Honeywell Labs, to build an \emph{architectural domain-specific language} (ADSL) for specifying and verifying distributed fault-tolerant systems. The ADSL should be able to synthesize both software and/or hardware implementations as well as formal models for verification. Before building such an ADSL, we needed a scalable general formal model to which to compile, leading to the work presented in this paper. We hypothesize that the ADSL will make refactoring even easier, and we can generate invariants or invariant templates useful for verification. Indeed, we have developed a preliminary ADSL that generates C code as well as formal models in SRI's Sally~\cite{pdr-kind}, to be described in future work.\footnote{\url{https://github.com/GaloisInc/atom-sally}}

>>>>

\usepackage[nocompress]{cite}
\usepackage{alltt}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{ifthen}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{tikz-cd}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{xspace}

@@

\author{Benjamin F Jones \and Lee Pike}

@@

\begin{abstract}
With proof techniques like IC3 and $k$-induction, model-checking scales further
than ever before. Still, fault-tolerant distributed systems are particularly
challenging to model-check given their large state spaces and
non-determinism. The typical approach to controlling complexity is to construct
ad-hoc abstractions of faults, message-passing, and behaviors. However, these
abstractions come at the price of divorcing the model from its implementation
and making refactoring difficult. In this work, we present a model for
fault-tolerant distributed system verification that combines ideas from the
literature including calendar automata, symbolic fault injection, and abstract
transition systems, and then use it to model-checking various implementations of
the Hybrid Oral Messages algorithm that differ in the fault model, timing model,
and local node behavior. We show that despite being implementation-level models,
the verifications are scalable and modular, insofar as isolated changes to an
implementation require isolated changes to the model and proofs. This work is
carried out in the SAL model-checker.
\end{abstract}

@@

In this paper, we present a fault-tolerant distributed systems model, and use that model to verify several variant implementations of the Byzantine fault-tolerant Hybrid Oral Messages algorithm (\OMH)~\cite{Lincoln-Rushby}. The model combines various ideas from the literature to build scalable and modular formal models suitable for infinite-state model-checking, and it reduces the need for ad-hoc abstractions and optimizations. In Section~\ref{sec:model}, we present the important aspects of the model, including calendar automata, originally developed by Dutertre and Sorea~\cite{cal}, symbolic fault-injection, and abstract transition systems for verification.

@@

Let a transition system $\mathcal{M} = (S, I, \rightarrow)$, be a set of states $S$, a set of initial states $I \subseteq S$, and a transition relation $\rightarrow \subseteq S \times S$. We implicitly assume a set of state variables such that each state $\sigma \in S$ is a total function that maps state variables to values. We sometimes prime a state to denote that it satisfies the transition relation: $\sigma \rightarrow \sigma'$. We also sometimes use a variable assignment notation to describe what state variables are specifically updated: e.g., $\sigma' = \sigma[v := v+1]$.

@@

\item \label{cal:d} In all states, if the current time equals a timeout, then the only transitions enabled are calendar event updates associated with the timeout: $\forall \sigma \in S$, $\exists (e_i, t_i) \in \sigma(cal)$ such that $\sigma(now) = t_i$ implies $\forall \sigma'$ such that $\sigma \rightarrow \sigma'$, $\sigma'(now) = \sigma(now)$, $\sigma'(c_j) = \sigma(c_j)$ for all $c_j \in \sigma(cal)$ such that $c_j \neq c_i$ (recalling that by convention, $c_i = (e_i, t_i)$), and $c_i \notin \sigma'(cal)$.

@@

The Oral Messages algorithm and its variants and its variants have a long history of formal verification. \OM{1} was verified in both the PVS and ACL2 interactive theorem-provers~\cite{Young97:IC}. Also in ACL2, an implementation of a circuit design to implement \OM{1} is given~\cite{om-acl2-impl}; the low-level model most closely relates to our level of detail. A refinement-based verification approach is used, and \OM{1} is specialized to a fixed number of nodes. Bokor~\emph{et~al.} describe a message-passing model for synchronous distributed algorithms that is particularly amenable to partial-order reduction for explicit-state model-checking~\cite{Bokor2010}. The model is efficient for up to five nodes, but results are not presented beyond that. Very recently, Jovanovi{\'{c}} and Dutertre use a ``flattened'' high-level model of \OM{1} as a benchmark for IC3 augmented with $k$-induction~\cite{pdr-kind}.

@@

% ------------------------------------------------------------
\section{Conclusions}\label{sec:conclusions}
This work fits within a larger project, in collaboration with Honeywell Labs, to build an \emph{architectural domain-specific language} (ADSL) for specifying and verifying distributed fault-tolerant systems. The ADSL should be able to synthesize both software and/or hardware implementations as well as formal models for verification. Before building such an ADSL, we needed a scalable general formal model to which to compile, leading to the work presented in this paper. We hypothesize that the ADSL will make refactoring even easier, and we can generate invariants or invariant templates useful for verification. Indeed, we have developed a preliminary ADSL that generates C code as well as formal models in SRI's Sally~\cite{pdr-kind}, to be described in a future paper.\footnote{\url{https://github.com/GaloisInc/atom-sally}}
